The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2012-12-07 script/en/chapters/introduction.tex
CT script: Minor changes (updated intro)

%This edition of the CT script again updated some topics:
This edition completely updated the TeX sources of the document, and of course
the content of the script was amended, corrected, and updated with some topics, e.g.:
\vspace{-5pt}
\begin{itemize}
  \item the definitions of the strength of security functions
        (chap. \ref{cm_Section_Security_Definitions}),
%  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
%        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies});
        and where primes are used as hangers  
        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}),
  \item the overviews of all functions in
        \hyperlink{appendix-template-overview-CT2}{for CT2} and
        \hyperlink{appendix-function-overview-JCT}{for JCT},
  \item further Sage scripts for cryptography, and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        Sage. Sage becomes more and more the standard open-source CAS system,
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
  \item the section about the Goldbach conjecture
        (see \ref{L-GoldbachConjecture}),
  \item the section about shared primes in RSA modules used in reality
        (see \ref{l:NumberTheory_Shared-Primesxxx}),xxxxxxxxxxxxxxxxxxxx
  \item the section about RSA fixed points
        (see \ref{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}), and
  \item the chapter about homomorphic encryption (see \ref{Chapter_HomomorphicEnc}).
\end{itemize}

The script was delivered the first time with the CrypTool-1 package in
version 1.2.01. Since then it has been expanded and revised in almost
every new version of CrypTool 1.

@@

I am deeply grateful to all the people helping with their impressive
commitment who have made this global project so successful.
Thanks also to the readers who sent us feedback.
% Especially I would like to acknowledge the English language proof-reading
% of this script version done by Richard Christensen and Lowell Montgomery.

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\\
\\
Bernhard Esslinger
\\
\\
Frankfurt and Siegen (Germany), October 2012

@@

This document is delivered together with the open-source program
CrypTool 1\index{CrypTool 1}.

@@

Chapters  \ref{Chapter_ModernCryptography} (Modern Cryptography) and 
\ref{Chapter_EllipticCurves} (Elliptic Curves) require a deeper knowledge
in mathematics, while the other chapters should be understandable with a 
school leaving certificate.

@@

\noindent \textbf{CrypTool-Online and MTC3\index{CrypTool-Online}\index{CTO}\index{MTC3}}

Also part of the CT project is the website CrypTool-Online (CTO)
(\url{http://www.cryptool-online.org}), where you can check and apply cryptographic
methods within a browser or on a smartphone. The scope of CTO is far below from
the standalone programs CT1, CT2 and JCT.

Also based on the CT project is the international cryptography contest
MysteryTwister C3 (MTC3) (\url{http://www.mysterytwisterc3.org}).
Here you can find cryptographic riddles in four categories, a
high-score list and a moderated forum. In the meantime more than 3300 active
users participate, and more than 140 challenges are offered (currently 89 of
them are solved by at least one user).
\par \vskip + 15pt

@@

Sage is an open source CAS package which can be used to easily program
the mathematical methods explained in this script.
\par \vskip + 15pt

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to the CrypTool projct\index{CrypTool}. They
applied their very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Hr.\ Dominik Schadow
   \item Staff of 
         Prof.\ Johannes Buchmann,
         Prof.\ Claudia Eckert,
         Prof.\ Alexander May,
         Prof.\ Torben Weis and
         Prof.\ Arno Wacker.
\end{list}
%\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\
\\
Bernhard Esslinger
\\
\\
Frankfurt and Siegen (Germany), October 2012
\par \vskip + 15pt

\noindent PS:\\
We'd be glad if further authors would show up to add further chapters, e.g. about
hash functions, random numbers or the design and attack of crypto protocols (like SSL).

>>>>

%This edition of the CT script again updated some topics:
This edition completely updated the TeX sources of the document, and of course
the content of the script was amended, corrected, and updated with some topics, e.g.:
\vspace{-5pt}
\begin{itemize}
  \item the definitions of the strength of security functions
        (chap. \ref{cm_Section_Security_Definitions}),
%  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
%        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies}),
%       and where primes are used as hangers  
%        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}),
  \item the overviews of all functions in
        \hyperlink{appendix-template-overview-CT2}{for CrypTool 2 (CT2)},
        \hyperlink{appendix-function-overview-JCT}{for JCrypTool (JCT)}, and
        \hyperlink{appendix-function-overview-CTO}{for CrypTool-Online (CTO)},
  \item further Sage scripts for cryptography, and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        Sage,
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
  \item the section about the Goldbach conjecture
        (see \ref{L-GoldbachConjecture}),
  \item the section about shared primes in RSA modules used in reality
        (see \ref{l:NumberTheory_Shared-Primesxxx}),xxxxxxxxxxxxxxxxxxxx
  \item the section about RSA fixed points
        (see \ref{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}), and
  \item the chapter about homomorphic encryption (see \ref{Chapter_HomomorphicEnc}).
\end{itemize}

The script was delivered the first time with the CT1 package in
version 1.2.01. Since then it has been expanded and revised in almost
every new version of CT1.

@@

\newpage
\noindent \textbf{Acknowledgment}

At this point I'd like to thank explicitly the following people who
particularly contributed to the CrypTool projct\index{CrypTool}. They
applied their very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Hr.\ Dominik Schadow
   \item Staff of 
         Prof.\ Johannes Buchmann,
         Prof.\ Claudia Eckert,
         Prof.\ Alexander May,
         Prof.\ Torben Weis and especially
         Prof.\ Arno Wacker.
\end{list}

Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).

Thanks also to the readers who sent us feedback.
% Especially I would like to acknowledge the English language proof-reading
% of this script version done by Richard Christensen and Lowell Montgomery.

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\par \vskip + 15pt

\noindent Bernhard Esslinger
\\
\\
Frankfurt and Siegen (Germany), December 2012
\par \vskip + 30pt

\noindent PS:\\
We'd be glad if further authors would show up to add further chapters,
e.g. about the Riemann Zeta function, hash functions, random numbers, or
the design and attack of crypto protocols (like SSL).

@@

This document is delivered together with the open-source programs of the
CrypTool project\index{CrypTool}.

@@

Chapters \ref{Chapter_ModernCryptography} (Modern Cryptography),
\ref{Chapter_EllipticCurves} (Elliptic Curves), and
\ref{Chapter_HomomorphicEnc}(Homomorphic Encryption)
require a deeper knowledge in mathematics, while the other chapters should
be understandable with a school leaving certificate.

@@

\noindent \textbf{CrypTool-Online\index{CrypTool-Online}\index{CTO}}

Also part of the CT project is the website CrypTool-Online (CTO)
(\url{http://www.cryptool-online.org}), where you can check and apply
cryptographic methods within a browser or on a smartphone. The scope
of CTO is far below from the standalone programs CT1, CT2 and JCT.
\par \vskip + 15pt

\noindent \textbf{MTC3\index{MTC3}}

MTC3 is the abbreviation for MysteryTwister C3, an international cryptography
contest (\url{http://www.mysterytwisterc3.org}), which is also based on the
CT project.
Here you can find cryptographic riddles in four categories, a
high-score list and a moderated forum. In the meantime more than 3500 active
users participate, and more than 140 challenges are offered (currently 89 of
them are solved by at least one user).
\par \vskip + 25pt  % more than +15 to start Sag at the next page

@@

Sage is a comprehensive open-source CAS package which can be used to easily
program the mathematical methods explained in this script. A speciality of
this CAS is, that the scripting language is Python.\\
Sage becomes more and more the standard open-source CAS system at universities.
\par \vskip + 15pt

\noindent \textbf{The Pupil's Crypto Courses\index{Pupil's Crypto}}

Within this initiative, one and two 2 day courses in cryptology are offered
for pupils and teachers in order to show how attractive MINT subjects like
mathematics, computer science and especially cryptology are.
The course agenda is a virtual secret agent story.\\
In the meantime, these courses took place for several years in Germany in
different towns.\\
All course material is freely available on net
(\url{http://www.cryptool.org/schuelerkrypto/}).\\
All software used is free software (using mostly CT1 and CT2).\\
As all course material is currently only available in German we'd be happy
if someone could translate the course material and build an according course
in English.
\par \vskip + 15pt

@@

I am deeply grateful to all the people helping with their impressive
commitment who have made this global project so successful.
\par \vskip + 15pt

\noindent Bernhard Esslinger
\\
\\
Frankfurt and Siegen (Germany), December 2012

### 2012-12-07 script/en/script-en.tex
CT script: Minor changes (updated intro)

% Hier wird \vspace{32pt} einfach ignoriert
\author{ % Die Klammer zur Fontänderung funktioniert bei author nur einzeilig im Gegensatz zu title?!
        {\LARGE Copyright (c) Prof.\ Bernhard Esslinger (co-author and editor)}\\[5pt]
        {\LARGE and the CrypTool Development Team, 1998-2012}\\[5pt]
        {\LARGE Frankfurt am Main, Germany} \\*[15pt]
        \url{www.cryptool.org}
}

@@

% Vergleichbar zur Nutzung von Texten aus der Wikipedia (=GNU Free Documentation Licence, GFDL)
% \vspace{200pt}
\begin{quote}
This is a free document, so the content of the document can
be copied and distributed, also for commercial purposes --- as long as
the authors, title and the CrypTool web site (\url{www.cryptool.org})
are acknowledged. Naturally, citations from the CrypTool script are
possible, as in all other documents.\\
The specific license for this document is the GNU Free Documentation Licence.

    Copyright \copyright{} 1998--2012 Bernhard Esslinger and the
    CrypTool Development Team. Permission is granted to copy,
    distribute and/or modify this document under the terms of the GNU
    Free Documentation License, Version 1.3 or any later version
    published by the Free Software Foundation; with no Invariant
    Sections, no Front-Cover Texts, and no Back-Cover Texts. A copy of
    the license is included in the section entitled
    \hyperlink{appendix-GNU-fdl}{``GNU Free Documentation License''}.
\end{quote}

\vspace{460pt}
Source cover photograph: \url{www.photocase.com}, Andre Guenther\\

>>>>

% Hier wird \vspace{32pt} einfach ignoriert
\author{ % Die Klammer zur Fontänderung funktioniert bei author nur einzeilig im Gegensatz zu title?!
        {\LARGE Copyright (c) Prof.\ Bernhard Esslinger (co-author and editor)}\\[5pt]
        {\LARGE and the CrypTool Team, 1998-2013}\\[5pt]
        {\LARGE Frankfurt am Main, Germany} \\*[15pt]
        \url{www.cryptool.org}
}

@@

% Vergleichbar zur Nutzung von Texten aus der Wikipedia (=GNU Free Documentation Licence, GFDL)
% \vspace{200pt}
\begin{quote}
This is a free document, so the content of the document can
be copied and distributed, also for commercial purposes --- as long as
the authors, title and the CrypTool web site (\url{www.cryptool.org})
are acknowledged. Naturally, citations from the CrypTool script are
possible, as in all other documents.\\
Additionally, this document is liable to the specific license of the
GNU Free Documentation Licence.

    Copyright \copyright{} 1998--2013 Bernhard Esslinger and the
    CrypTool Team. Permission is granted to copy,
    distribute and/or modify this document under the terms of the GNU
    Free Documentation License, Version 1.3 or any later version
    published by the Free Software Foundation. A copy of
    the license is included in the section entitled
    \hyperlink{appendix-GNU-fdl}{``GNU Free Documentation License''}.

    This also includes the code of the Sage samples in this document.
\end{quote}

\vspace{420pt}
Source cover photograph: \url{www.photocase.com}, Andre Guenther\\
